---
title: "Characterising prevalence and most common sets of multiple long terms conditions"
format: 
    html: yes
    toc: yes
    toc_float: yes
    number_sections: yes
    toc_depth: 5
editor: visual
---

# Objective

\|\> To describe the prevalence of multimorbidity in IPCI and identify the clusters of associated diseases.

# Methods

## Study design:

-   Retrospective cohort study.

## Setting:

-   IPCI database. Patient records from 350 General Practioners (GPs) in the Netherlands.

## Participants.

Using the OMOP CDM,\
- **Cohort definition**: "All patients in the database, above 18 years old, with an observation period starting any time before 31st December 2022."\
- Information for the participants will be extracted from the database until 31st December 2022. Index date will be the start of the observation period for each patient.\
- I will also consider as a requirement the presence of at least a year of prior observation as additional eligibility criteria. This was not implement for the results in this report.

## Variables:

-   We extract all information which comprises of 75 conditions that can be used to define multimorbidity according to Ngoran et al (2016).\
-   We define multimorbidity as the presence of two or more conditions, out of the list of 75 conditions.\
-   Only the first occurrence of a disease will be included. For example, if a patient experienced stroke on two separate occasions, only the first will be considered an indication that the person is a stroke patient.

## Statistical methods:

1)  Measure prevalence of multimorbidity by simple counting.\

-   Stratification by age and age groups.\

2)  Association rule mining to identify the clusters of associative diseases.\
3)  Sensitivity analyses:\

-   Possibility to explore other definitions. Other definitions include, usually, a differing number of diseases (cite Ho, Delphi study).\

4)  Potential sources of bias.

-   Data linkage (same patient be represented by multiple observation periods).
-   Use of concepts instead of phenotype definitions.
-   Misclassification of conditions.
-   Varied follow-up times.

# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(tidyverse)
library(arules)
library(Andromeda)
library(FeatureExtraction)
library(ggplot2)
library(tableone)
library(knitr)
library(kableExtra)
library(DT)
library(magrittr)
library(gt)
library(ggpol)
library(gridExtra)
library(arulesViz)
library(plotly)
library(igraph)
library(ggraph)
options(scipen = 999)
```

```{r, include=FALSE}
createMultimorbidityDataframe <- function(covariateData){
  
  ageDf <- covariateData$covariates %>%
    dplyr::filter(covariateId  == 1002) %>%
    arrow::to_duckdb() %>%
    dplyr::mutate(covariateId = as.character(covariateId)) %>%
    tidyr::pivot_wider(id_cols = rowId, names_from = covariateId, values_from = covariateValue) %>%
    dplyr::rename(age = `1002.0`) %>%
    arrow::to_arrow() %>%
    dplyr::collect()
  
  genderDf <- covariateData$covariates %>%
    dplyr::filter(covariateId %in% c(8507001, 8532001)) %>%
    arrow::to_duckdb() %>%
    dplyr::mutate(gender = case_when(
      covariateId == 8507001 ~ "Male", 
      covariateId == 8532001 ~ "Female"
    )) %>%
    dplyr::select(rowId, gender) %>%
    arrow::to_arrow() %>%
    dplyr::collect()
  
  if (FeatureExtraction::isTemporalCovariateData(covariateData)){
    conditionsDf <- covariateData$covariates %>%
      dplyr::filter(!(covariateId %in% c(1002, 8507001, 8532001))) %>%
      arrow::to_duckdb() %>%
      dplyr::group_by(rowId, covariateId) %>%
      dplyr::slice_max(timeId, with_ties = FALSE) %>%
      dplyr::ungroup() %>%
      dplyr::collect()
  } else {
    conditionsDf <- covariateData$covariates %>%
      dplyr::filter(!(covariateId %in% c(1002, 8507001, 8532001))) %>%
      dplyr::collect() 
  }
  
  
  multimorbidityDf <- ageDf %>%  
    inner_join(genderDf, by = "rowId") %>%
    full_join(conditionsDf, by = "rowId", multiple = "all") %>%
    # arrow::to_duckdb() %>%
    mutate(covariateValue = case_when(
      is.na(covariateId) ~ "", 
      .default = as.character(covariateValue)
    )) %>%
    # arrow::to_arrow() %>%
    full_join(covariateData$covariateRef %>% filter(!(covariateId %in% c(1002, 8507001, 8532001))) %>% collect(), by = "covariateId") %>%
    arrange(rowId) %>% collect()
  
  return(multimorbidityDf)
}
```

```{r, include=FALSE}
covariatesTemp <- FeatureExtraction::loadCovariateData("experimentation/Multimorbidity_20240110/tempCovs_cdm_O_selectedConceptIds_obsrvtn_4")

```

```{r, include=FALSE}
inputTemp <- createMultimorbidityDataframe(covariateData = covariatesTemp)
```

```{r}
inputTemp %>% group_by(covariateName) %>% distinct()
length(unique(inputTemp$covariateName))
# unique(inputTemp$covariateName)
length(unique(inputTemp$covariateId))
```

```{r}
Ho_conditions <- read.csv("final_CodeListToInclude.csv")
```

```{r}
dbCovIds <- inputTemp%>% distinct(covariateId) %>%
  mutate(newCovId = as.numeric(gsub('.{3}$', '', covariateId)))
HoCovIds <- Ho_conditions %>% distinct(newSnomedCodes)

# Codes included 
HoCovIds$newSnomedCodes[HoCovIds$newSnomedCodes%in% dbCovIds$newCovId]
# Codes not included
notIncluded <- HoCovIds$newSnomedCodes[!(HoCovIds$newSnomedCodes%in% dbCovIds$newCovId)]

notInDf <- Ho_conditions %>% filter(newSnomedCodes %in% notIncluded)

covTemp2 <- FeatureExtraction::loadCovariateData("experimentation/Multimorbidity_20240110_old/tempCovs_cdm_O_selectedConceptIds_obsrvtn_4")
inputTemp2 <- createMultimorbidityDataframe(covariateData = covariatesTemp)
dbCovIds2 <- inputTemp2 %>% distinct(covariateId) %>%
  mutate(newCovId = as.numeric(gsub('.{3}$', '', covariateId)))
notIncluded2 <- HoCovIds$newSnomedCodes[!(HoCovIds$newSnomedCodes%in% dbCovIds2$newCovId)]
```

```{r}

Ho_conditions$newSnomedNames %in% unique(inputTemp$covariateName)


inputTemp%>% distinct(covariateId) %>% count()
length(Ho_conditions$newSnomedCodes)
length(unique(Ho_conditions$newSnomedCodes))
Ho_conditions %>% group_by(newSnomedCodes) %>% count() %>% filter(n>1)

concepts <- covariatesTemp$covariateRef %>%
  collect() %>%
  left_join(
    Ho_conditions %>%
      select(Include, Body.system..based.on.ICD.10.chapters., Ho_Name, newSnomedCodes) %>%
      rename("Body system" = "Body.system..based.on.ICD.10.chapters."), 
    join_by(conceptId == newSnomedCodes)
  )

df1 <- inputTemp %>%
  left_join(
    concepts %>% select(conceptId, Ho_Name, `Body system`), 
    by = "conceptId", 
    na_matches = 'never'
  )

```

```{r}
# Merging concepts from Ho et al

concepts_2 <- concepts %>%
  mutate(Ho_et_al_conditions = case_when(
    str_detect(condition, "Malignan|lymphoma") ~ "Cancer", 
    str_detect(condition, "pain|Pain") ~ "Chronic pain",
    str_detect(condition, "dementia|Dementia") ~ "Dementia",
    str_detect(condition, "Depress|depress") ~ "Depression",
    str_detect(condition, "Drug|drug|Tobacco|Medication abuse") ~ "Drug use disorders", 
    str_detect(condition, "Alcohol|alcohol") ~ "Alcohol use disorders", 
    str_detect(condition, "Asthma|asthma") ~ "Asthma", 
    str_detect(condition, "Chronic obstructive") ~ "COPD", 
    str_detect(condition, "Schizophrenia|psychosis|Psychosis") ~ "Schizophrenia", 
    str_detect(sourceName, "1 diabetes") ~ "T1DM", 
    str_detect(sourceName, "2 diabetes") ~ "T2DM", 
    str_detect(condition, "Osteoporosis") ~ "Osteoporosis", 
    str_detect(condition, "arthritis|Arthritis|Gout") ~ "Arthritis",
    str_detect(condition, "Retino|retino|Blind|blind|Macular") ~ "Vision impairment",
    str_detect(condition, "Hear|Deaf") ~ "Hear impairment",
    str_detect(condition, "of trauma") ~ "Musculoskeletal impairment due to injury", 
    str_detect(Category, "Cardiovascular") ~ "Coronary heart disease",
    str_detect(condition, "Cerebro") ~ "Stroke",
    str_detect(Category, "Neurological") ~ "Neurological", 
    str_detect(condition, "HIV") ~ "HIV", 
    str_detect(condition, "bowel syndrome|colitis") ~ "IBD", 
    str_detect(condition, "stress") ~ "Anxiety",
    Category ==  "Psychological" & str_detect(condition, "Schizophrenia|psychosis|Psychosis|Depress|depress|stress", negate =TRUE) ~ "Other psychological", 
    Category == "Skin" ~ "Skin ulcers", 
    str_detect(condition, "Anorexia") ~ "Anorexia/Bulimia",
    str_detect(condition, "Obesity") ~ "Obesity",
    str_detect(condition, "Incontinence Urine") ~ "Incontinence Urine",
    str_detect(condition, "Chronic bronchitis") ~ "Chronic bronchitis", 
    .default = NA
  ))

concepts_2 <- concepts_2 %>%
  mutate(Ho_and_ICPC = case_when(
    is.na(Ho_et_al_conditions) ~ condition, 
    .default = Ho_et_al_conditions
  ))

concepts <- concepts_2 %>%
  dplyr::rename(
    "Snomed_condition" = "condition", 
    "condition" = "newCondNames")
```

```{r, include=FALSE}
inputTemp2 <- concepts %>%
  select(conceptId, refined_ICPCcode, conceptName, condition) %>%
  inner_join(inputTemp, join_by(conceptId == conceptId), multiple = "all") %>%
  group_by(rowId, condition) %>%
  slice(1L) %>%
  ungroup()

######

df_inputTemp <- as.data.frame(dplyr::select(inputTemp2, c(rowId, condition)))
str(df_inputTemp)

noDiseaseTemp <- inputTemp %>%
  filter(is.na(conceptId)) %>%
  select(rowId, covariateName)

df_inputTemp2 <- bind_rows(df_inputTemp, noDiseaseTemp) %>% select(rowId, condition)
```

## Methods

```{r, include=FALSE}
# c1 <- concepts %>% 
#   select(refined_ICPCcode, condition, Category) %>%
#   distinct() %>%
#   DT::datatable(caption = "List of conditions included in the study.")

c1 <- concepts %>% 
  select(refined_ICPCcode, condition, Category) %>%
  distinct() %>%
  rename("ICPC-1 Code" = "refined_ICPCcode") %>%
  gt() %>%
  tab_caption(caption = "Table S2: Conditions included in the definition of multimorbidity.")

c2 <- concepts %>% 
  select(refined_ICPCcode, condition, Category) %>%
  distinct() %>%
  group_by(Category) %>%
  count() %>%
  ungroup() %>%
  rename("Number of Conditions" = "n") %>%
  gt() %>%
  tab_caption(caption = "Table S1: Categorisation of diseases included in the definition of Multimorbidity.")
```

### Categorisation of conditions included.

Table S1: Categorisation of diseases included in the definition of Multimorbidity.\
`r c2`

------------------------------------------------------------------------

### Detailed conditions used.

Table S2: Conditions included in the definition of multimorbidity.\
`r c1`

------------------------------------------------------------------------

```{r, include=FALSE}
# Some numbers
population <- length(unique(df_inputTemp2$rowId))
populationWithCondition <- df_inputTemp2 %>%
  filter(condition != "") %>%
  distinct(rowId) %>%
  count() %>%
  pull()

populationWithoutConditionDf <- df_inputTemp2 %>%
  filter(condition == "") %>%
  distinct(rowId) %>%
  count() %>%
  rename("NoPeople"= "n") %>%
  mutate(n = 0)

populationWithoutCondition <- population -populationWithCondition

peopleWithoutConditionDf <- df_inputTemp2 %>%
  filter(is.na(condition)) %>%
  select(rowId, condition) %>%
  rename("n" = "condition") %>%
  mutate(n = 0)

## how many people have what condition
numberOfCOnditions <- df_inputTemp2 %>%
  filter(condition != "") %>%
  group_by(rowId) %>%
  summarise(n = n()) %>%
  group_by(n) %>%
  summarise(NoPeople = n(), 
            Perc = (NoPeople/population)*100, 
            Perc = round(Perc, 5), 
            PercAmongDisease = (NoPeople/populationWithCondition)*100, 
            PercAmongDisease = round(PercAmongDisease, 5))
sum(numberOfCOnditions$NoPeople) == populationWithCondition

numberOfCOnditions %>%
  filter(n != 1) %>%
  summarise(NoPeople = sum(NoPeople), 
            Perc = sum(Perc), 
            PercAmongDisease = sum(PercAmongDisease))

numberOfCOnditions2 <- df_inputTemp2 %>%
  filter(!is.na(condition)) %>%
  group_by(rowId) %>%
  summarise(n = n()) %>%
  bind_rows(peopleWithoutConditionDf) %>%
  group_by(n) %>%
  summarise(NoPeople = n()) 

rowIdPeopleWithMultimorbidity <- df_inputTemp2 %>%
  filter(!is.na(condition)) %>%
  group_by(rowId) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n >=2)

nrow(rowIdPeopleWithMultimorbidity)/population

peopleWithMultimorbidityDF <- inner_join(rowIdPeopleWithMultimorbidity, df_inputTemp2, by = "rowId") 
length(unique(peopleWithMultimorbidityDF$rowId))

peopleWithMultimorbidityDF %>%
  group_by(condition) %>%
  summarise(n = n()) %>%
  mutate(perc = (n/length(unique(peopleWithMultimorbidityDF$rowId))) * 100 ) %>%
  arrange(desc(perc))

gdf <- inputTemp %>% 
  filter(rowId %in% peopleWithMultimorbidityDF$rowId) %>%
  select(rowId, gender) %>%
  distinct() %>%
  inner_join(peopleWithMultimorbidityDF, by = "rowId", multiple = "all")

gdf_Time <- gdf %>%
  inner_join(inputTemp2 %>% select(rowId, condition, age, timeId) %>% distinct(), by = join_by(rowId == rowId, condition == condition), multiple = "all")

baselineDf1 <- gdf_Time %>%
  filter(timeId == 0) %>%
  group_by(rowId) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  mutate(multimorbid = case_when(
           n >=2 ~ "Yes", 
           .default = "No"
         )) 

baselineDf <- inputTemp %>% 
  select(rowId, age, gender) %>%
  distinct() %>%
  mutate(multimorbid = case_when(
    rowId %in% baselineDf1$rowId ~ "Yes", 
    .default = "No"
  ))

malePopulation <- baselineDf %>% 
  select(rowId, gender) %>%
  filter(gender == "Male") %>%
  nrow() 

femalePopulation <- baselineDf %>% 
  select(rowId, gender) %>%
  filter(gender == "Female") %>%
  nrow() 
```

## Results

### Baseline Characteristics

```{r, include=FALSE}
# inputTemp %>%
#   select(rowId, age, gender) %>%
#   distinct() %>% 
#   summarise(N = n(), 
#             N_mean_age = mean(age)) %>%
#   bind_rows(inputTemp %>%
#   select(rowId, age, gender) %>%
#   distinct() %>%
#     group_by(gender) %>%
#     summarise(N_g = n(), 
#               age_g = mean(age)))
# 
# inputTemp %>%
#   select(rowId, age, gender) %>%
#   distinct() %>% 
#   filter(gender == "Male") %>%
#   summarise(mean_age = as.numeric(sum(age)/n()))


t1 <- tableone::CreateTableOne(data = baselineDf, vars = c("age", "multimorbid"), factorVars = c("multimorbid"), addOverall = TRUE, strata = "gender")
t1

t1_gt <- as.data.frame(print(t1, quote = FALSE, noSpaces = TRUE, printToggle = FALSE, formatOptions = list(big.mark = ","))) %>%
  mutate(name = rownames(.)) %>%
  select(name, 1:4) %>%
  rename(" " = "name") %>%
  gt() %>%
  tab_caption(caption = "Table 1: Baseline characteristics.")
  
```

Table 1: Baseline characteristics.\
`r t1_gt`

```{r, include=FALSE}
inputTemp2 %>% 
  select(rowId, gender, condition) %>%
  inner_join(baselineDf %>% select(rowId, multimorbid), by = "rowId") %>%
  group_by(condition, multimorbid) %>%
  count() %>%
  pivot_wider(id_cols = condition, names_from = multimorbid, values_from = n)

t2 <- inputTemp2 %>% 
  select(rowId, gender, condition) %>%
  inner_join(baselineDf %>% select(rowId, multimorbid), by = "rowId") %>%
  group_by(condition) %>%
  summarise(n = n()) %>%
  inner_join(
    inputTemp2 %>% 
  select(rowId, gender, condition) %>%
  inner_join(baselineDf %>% select(rowId, multimorbid), by = "rowId") %>%
  group_by(condition, multimorbid) %>%
  count() %>%
  ungroup()%>%
  pivot_wider(id_cols = condition, names_from = multimorbid, values_from = n), 
  by = "condition"
  ) %>%
  mutate(overall_perc = round((n/population)* 100, 2), 
         No_perc = round((No / populationWithoutCondition) * 100, 2), 
         Yes_perc = round((Yes / nrow(rowIdPeopleWithMultimorbidity))*100, 2)) %>%
  select(condition, n, overall_perc, No, No_perc, Yes, Yes_perc) %>%
  mutate("N (%)" = paste0(format(n, big.mark=","), " (", overall_perc, ")"), 
         "Not Multimorbid (%)" = paste0(format(No, big.mark=","), " (", No_perc, ")"), 
         "Multimorbid (%)" = paste0(format(Yes, big.mark=","), " (", Yes_perc, ")")) %>%
  select(condition, ends_with("(%)")) %>%
  rename("Condition" = "condition") 
  # kable() %>%
  # add_header_above(c(" ", "Overall" = 2, "Multimorbid" = 4))

```

Table S3: Prevalence of each condition at baseline.\
`r t2 %>% gt()`

------------------------------------------------------------------------

### Prevalence

#### Overall Prevalence of Conditions and stratified by gender

```{r, include=FALSE}
numberOfCOnditions22 <- df_inputTemp %>%
  group_by(rowId) %>%
  summarise(n = n()) %>%
  bind_rows(peopleWithoutConditionDf) %>%
  group_by(n) %>%
  summarise(NoPeople = n())
# 
# numberOfCOnditions2 <- df_inputTemp2 %>%
#   group_by(rowId) %>%
#   summarise(n = n()) %>%
#   ungroup() %>%
#   group_by(n) %>%
#   summarise(NoPeople = n())

# saveRDS(numberOfCOnditions22, "experimentation/results_presentation/numberOfConditions22.Rds")

# numberOfCOnditions22 <- readRDS("results_presentation/numberOfConditions22.Rds")
tableForNumberOfConditions <- numberOfCOnditions22 %>%
  mutate(n = as.character(n)) %>%
  add_row(n = "Eight or More", NoPeople = numberOfCOnditions22 %>% dplyr::filter(n>= 8) %>% summarise(new = sum(NoPeople)) %>% pull()) %>%
  mutate(Perc = (NoPeople/population)*100, 
            Perc = round(Perc, 2) 
            # PercAmongDisease = (NoPeople/populationWithCondition)*100, 
            # PercAmongDisease = round(PercAmongDisease, 5)
         ) %>%
 filter(!(n %in% c(8:30))) 
```

```{r, include=FALSE}
# tableForNumberOfConditions %>% gt() %>% tab_caption(caption = "Table 2: Prevalence of number of conditions in IPCI")
```

```{r, include=FALSE}
numberOfCOnditions4 <- inputTemp%>%
  # inner_join(baselineDf %>% select(rowId, gender), by = "rowId") %>%
  mutate(condPres = case_when(
    is.na(covariateId) ~ 0, 
    .default = 1
  )) %>%
  select(rowId, gender, condPres) %>%
  group_by(rowId, gender) %>%
  summarise(n = sum(condPres)) %>%
  ungroup() %>%
  # mutate(n = as.character(n)) %>%
  mutate(NumberOfConditions = case_when(
    n == 1 ~ "1",
    n == 2 ~ "2",
    n == 3 ~ "3",
    n == 4 ~ "4",
    n == 5 ~ "5",
    n == 6 ~ "6",
    n == 7 ~ "7",
    n >= 8 ~ "Eight or More", 
    .default = "0"))
# saveRDS(numberOfCOnditions4, "experimentation/results_presentation/numberOfConditions4.Rds")

# numberOfCOnditions4 <- readRDS("results_presentation/numberOfConditions4.Rds")

# t3 <- CreateTableOne(data = numberOfCOnditions4, vars = c("NumberOfConditions"), strata = "gender")
# 
# t3_gt <- as.data.frame(print(t3, quote = FALSE, noSpaces = TRUE, printToggle = FALSE, formatOptions = list(big.mark = ","))) %>%
#   mutate(name = rownames(.)) %>%
#   select(name, 1:2)  

t2new <- numberOfCOnditions4 %>% group_by(gender, NumberOfConditions) %>%
  summarise(n = n()) %>%
  pivot_wider(id_cols = NumberOfConditions, names_from = gender, values_from = n) %>%
  left_join(tableForNumberOfConditions, by = join_by("NumberOfConditions"=="n")) %>%
  mutate(Fem_perc = round(Female/population *100, 2 ), 
         Mal_perc = round(Male/population * 100, 2), 
         `Female (%)` = paste0(format(Female, big.mark=","), " (", Fem_perc, ")"), 
         `Male (%)` = paste0(format(Male, big.mark=","), " (", Mal_perc, ")"), 
         Overall = paste0(format(NoPeople, big.mark=","), " (", Perc, ")")
         ) %>%
  select(NumberOfConditions, Overall, ends_with("(%)"))
```

Table 2: Prevalence of number of conditions recorded for each patient. `r t2new %>% gt() %>% tab_caption(caption = "Table 2: Prevalence of number of conditions in IPCI")`

------------------------------------------------------------------------

```{r, include=FALSE}
numberOfCOnditions3 <- df_inputTemp2 %>%
  inner_join(baselineDf %>% select(rowId, gender), by = "rowId") %>%
  group_by(rowId, gender) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(n, gender) %>%
  summarise(NoPeople = n())

# saveRDS(numberOfCOnditions3, "experimentation/results_presentation/numberOfConditions3.Rds")
 
# numberOfCOnditions3 <- readRDS("results_presentation/numberOfConditions3.Rds")

nc3 <- numberOfCOnditions3 %>% 
  pivot_wider(id_cols = "n", names_from = gender, values_from = NoPeople) %>%
  filter(n != 1) 

t4 <-   tibble(Females = sum(nc3$Female), 
            Males = sum(nc3$Male, na.rm = TRUE)) %>%
    mutate(F_perc = (Females / population )* 100, 
            M_perc = (Males / population )* 100) %>%
  rename("Female (%)" = "F_perc", 
         "Male (%)" = "M_perc") %>%
  mutate(`% F among MM` = (Females / (Females + Males)) * 100, 
         `% M among MM` = (Males / (Females + Males)) * 100)

populationWithMultimorbidity <- t4$Females + t4$Males
# t4$Females/populationWithMultimorbidity *100
```

------------------------------------------------------------------------

Overall, prevalence of multimorbidity in IPCI using N'Goran's definition is `r format(populationWithMultimorbidity, big.mark=",")`, acounting for `r round((populationWithMultimorbidity)/ population *100, 2)`% of the total population.

------------------------------------------------------------------------

```{r, include=FALSE}
# Patients with multimorbidity
# gdf %>%
#   select(rowId, gender) %>%
#   distinct() %>%
#   group_by(gender) %>%
#   summarise(n = n(), 
#             perc = (n/length(unique(peopleWithMultimorbidityDF$rowId)))*100)
# 
# gdf_Time %>%
#   group_by(rowId) %>%
#   slice(1L) %>%
#   ungroup %>%
#   summarise(mean_age = mean(age))
# 
# gdf_Time %>% 
#   select(rowId, gender, age) %>%
#   distinct() %>%
#   summarise(mean_age = mean(age))
# 
# gdf_Time %>% 
#   select(rowId, gender, age) %>%
#   distinct() %>%
#   group_by(gender) %>%
#   summarise(mean_age = mean(age))

t5 <- tibble(Gender = c("Female", "Male"), 
             N = c(sum(nc3$Female), sum(nc3$Male, na.rm = TRUE))) %>%
  inner_join(gdf_Time %>% 
  select(rowId, gender, age) %>%
  distinct() %>%
  group_by(gender) %>%
  summarise(mean_age = mean(age), 
            sd_age = sd(age)), by = join_by("Gender" == "gender")) %>%
  mutate(`N (%)` = (N/population)*100, 
         `% among MM` = (N/populationWithMultimorbidity)*100, 
         `% among F` = (N/femalePopulation)* 100, 
         `% among M` = (N/malePopulation)* 100, 
         `% among F` = case_when(
           Gender == "Male" ~ 0, 
           .default = `% among F`
         ),
         `% among M` = case_when(
           Gender == "Female" ~ 0, 
           .default = `% among M`
         ), 
         N = format(N, big.mark=",")
         )
```

Table 3: Characterization of multimorbid population. `r t5 %>% gt() %>% tab_caption(caption = "Characterization of multimorbid population.")`

------------------------------------------------------------------------

### Multimorbid Conditions

```{r include=FALSE}
t6.0 <- gdf_Time %>%
  select(rowId, condition) %>%
  group_by(condition) %>%
  summarise(Overall = n()) %>%
  mutate(Overall_perc = round(Overall / populationWithMultimorbidity * 100, 2), 
         "Overall (%)" = paste0(format(Overall, big.mark = ","), " (", Overall_perc, ")")) %>%
  select(condition, ends_with("(%)"))

t6.1 <- gdf_Time %>% 
  select(rowId, gender, condition) %>%
  group_by(gender, condition) %>%
  summarise(N = n()) %>%
  pivot_wider(id_cols = condition, names_from = gender, values_from = N) %>%
  mutate(`F_perc` = round((Female/populationWithMultimorbidity )* 100, 2), 
         `M_perc` = round((Male/populationWithMultimorbidity )* 100, 2), 
         "Female (%)" = paste0(format(Female, big.mark=","), " (", F_perc, ")"), 
         "Male (%)" = paste0(format(Male, big.mark=","), " (", M_perc, ")")) %>%
  select(condition, ends_with("(%)")) %>%
  rename("Condition" = "condition")
  # DT::datatable(caption = "Prevalence of Individual Conditions within Multimorbid Patients") 

t6 <- right_join(t6.1, t6.0, by = join_by("Condition"== "condition")) %>%
  select(Condition, "Overall (%)", everything())

```

#### Prevalence of Conditions in multimorbid population

Table 4: Prevalence of conditions with the multimorbid population.

`r t6 %>% gt() %>% tab_caption(caption = "Prevalence of conditions with the multimorbid population.")`

```{r, include=FALSE}
mm_df <- inputTemp %>%
  group_by(rowId) %>%
  mutate(NoConditions = n())%>%
  ungroup() %>%
  arrange(rowId, timeId) %>%
  mutate(conditionAge = case_when(
    NoConditions >= 2 ~ age + round(-timeId/365), 
    .default = NA
  ))

mm_df2 <- mm_df %>%
  filter(NoConditions >=2) %>%
  arrange(rowId, conditionAge)

mm_df2 %>% distinct(rowId) %>% count
populationWithMultimorbidity

mm_df2 <- mm_df2 %>%
  group_by(rowId) %>%
  mutate(multimorbidAge = conditionAge[2]) %>%
  ungroup()
  

pretty_symmetric <- function(range, n = 5){
  range_1 <- c(-range[1], range[2])
  range_2 <- c(range[1], -range[2])
  
  pretty_vec_1 <- pretty(range_1)
  pretty_vec_2 <- pretty(range_2)
  
  pretty(
    c(pretty_vec_1, pretty_vec_2), 
    n = n
  )
}  

mm_df3 <- mm_df2 %>%
  group_by(rowId) %>%
  mutate(lastConditionAge = last(conditionAge)) %>%
  ungroup()

mm_df4 <- mm_df3 %>%
  group_by(rowId) %>%
  mutate(maxConditionAge = max(conditionAge)) %>%
  ungroup()

all.equal(mm_df4$lastConditionAge, mm_df4$maxConditionAge)
```

```{r, include=FALSE}
newPlotDf5 <- mm_df4 %>%
  group_by(rowId) %>%
  slice(1L) %>%
  ungroup() %>%
  mutate(ageGroup = case_when(
    maxConditionAge < 25 ~ "18-24", 
    maxConditionAge >= 25 & maxConditionAge < 30 ~ "25-29", 
    maxConditionAge >= 30 & maxConditionAge < 35 ~ "30-34", 
    maxConditionAge >= 35 & maxConditionAge < 40 ~ "35-39", 
    maxConditionAge >= 40 & maxConditionAge < 45 ~ "40-44",  
    maxConditionAge >= 45 & maxConditionAge < 50 ~ "45-49", 
    maxConditionAge >= 50 & maxConditionAge < 55 ~ "50-54", 
    maxConditionAge >= 55 & maxConditionAge < 60 ~ "55-59", 
    maxConditionAge >= 60 & maxConditionAge < 65 ~ "60-64",
    maxConditionAge >= 65 & maxConditionAge < 70 ~ "65-69", 
    maxConditionAge >= 70 & maxConditionAge < 75 ~ "70-74",
    maxConditionAge >= 75 & maxConditionAge < 80 ~ "75-79", 
    maxConditionAge >= 80 & maxConditionAge < 85 ~ "80-84",
    maxConditionAge >= 85 & maxConditionAge < 90 ~ "85-89", 
    maxConditionAge >= 90 & maxConditionAge < 95 ~ "90-94",
    maxConditionAge >= 95 & maxConditionAge < 100 ~ "95-99", 
    maxConditionAge >= 100 ~ "100+"  
  ), 
  groupedConditions = case_when(
    NoConditions >= 8 ~ "8+", 
    .default = as.character(NoConditions)
  ), 
  ageGroup = fct_relevel(ageGroup, "100+", after = Inf)) %>%
  # mutate(N = case_when(
  #   gender == "Male" ~ -1*N, 
  #   .default = N))
    # %>%
  select(rowId, gender, ageGroup, groupedConditions)

newPlotDf6 <- newPlotDf5 %>%
  group_by(gender, ageGroup, groupedConditions) %>%
  summarize(N = n())

# Figure 2: Multimorbid patients per age group stratified by gender 
p2 <- ggplot()+
  geom_bar(data = newPlotDf6[newPlotDf6$gender=="Male",], aes(x = ageGroup, y = N, fill = groupedConditions), stat = "identity",  alpha = 0.8, position = position_stack(reverse = TRUE))+
  geom_bar(data = newPlotDf6[newPlotDf6$gender=="Female",], aes(x = ageGroup, y = -N, fill = groupedConditions), stat = "identity", alpha = 0.8, position = position_stack(reverse = TRUE)) +
  facet_share(~gender, dir = "h",reverse_num = TRUE, scales = "free", switch = "both", strip.position = "top") + coord_flip()+
    labs(subtitle = "N = 553,262",
   caption = "NOTE: \n Patient's age at the last recorded condition was used to stratify patients in each age group.\n
   Figure 2: Multimorbid patients per age group stratified by gender.", 
   fill = "No of Conditions")+ 
  scale_color_manual(name = "", values = c(Male = "#3E606F", Female = "#8C3F4D"),
                 # labels = c("Females", "Males")
                 )+
  theme_classic() +
  theme(text = element_text(color = "#3A3F4A"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.2, color = "#3A3F4A"),
    # axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.x = element_text(size = 12, color = "#5D646F"),
    # axis.text.y = element_blank(),
    # strip.text = element_text(color = "#5D646F", size = 18, face = "bold", hjust = 0.030),
    plot.background = element_rect(fill = "#EFF2F4"),
    # plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "bottom",
    legend.spacing = unit(0.1, "lines"),
    legend.text  = element_text(size = 8),
    legend.text.align = 0, 
    legend.direction = "horizontal"
    ) +
  guides(fill = guide_legend(nrow = 1))
```

```{r, include=FALSE}
newPlotDf3 <- mm_df2 %>%
  group_by(rowId) %>%
  slice(1L) %>%
  ungroup() %>%
  mutate(ageGroup = case_when(
    multimorbidAge < 25 ~ "18-24", 
    multimorbidAge >= 25 & multimorbidAge < 30 ~ "25-29", 
    multimorbidAge >= 30 & multimorbidAge < 35 ~ "30-34", 
    multimorbidAge >= 35 & multimorbidAge < 40 ~ "35-39", 
    multimorbidAge >= 40 & multimorbidAge < 45 ~ "40-44",  
    multimorbidAge >= 45 & multimorbidAge < 50 ~ "45-49", 
    multimorbidAge >= 50 & multimorbidAge < 55 ~ "50-54", 
    multimorbidAge >= 55 & multimorbidAge < 60 ~ "55-59", 
    multimorbidAge >= 60 & multimorbidAge < 65 ~ "60-64",
    multimorbidAge >= 65 & multimorbidAge < 70 ~ "65-69", 
    multimorbidAge >= 70 & multimorbidAge < 75 ~ "70-74",
    multimorbidAge >= 75 & multimorbidAge < 80 ~ "75-79", 
    multimorbidAge >= 80 & multimorbidAge < 85 ~ "80-84",
    multimorbidAge >= 85 & multimorbidAge < 90 ~ "85-89", 
    multimorbidAge >= 90 & multimorbidAge < 95 ~ "90-94",
    multimorbidAge >= 95 & multimorbidAge < 100 ~ "95-99", 
    multimorbidAge >= 100 ~ "100+"  
  ), 
  groupedConditions = case_when(
    NoConditions >= 8 ~ "8+", 
    .default = as.character(NoConditions)
  ), 
  ageGroup = fct_relevel(ageGroup, "100+", after = Inf)) %>%
  # mutate(N = case_when(
  #   gender == "Male" ~ -1*N, 
  #   .default = N))
    # %>%
  select(rowId, gender, ageGroup, groupedConditions)

newPlotDf4 <- newPlotDf3 %>%
  group_by(gender, ageGroup, groupedConditions) %>%
  summarize(N = n())
newPlotDf7 <- newPlotDf4 %>%
  group_by(gender, ageGroup) %>%
  summarise(N = sum(N))

p1 <- ggplot(data = newPlotDf7,  aes(x = ageGroup, y = N, fill = gender))+
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_fill_manual(name = "", values = c(Male = "#3E606F", Female = "#8C3F4D"))+
  # geom_bar(data = newPlotDf4[newPlotDf4$gender=="Male",], aes(x = ageGroup, y = N), stat = "identity",  alpha = 0.8, position = position_stack(reverse = TRUE), fill =  "#3E606F")+
  # geom_bar(data = newPlotDf4[newPlotDf4$gender=="Female",], aes(x = ageGroup, y = N), stat = "identity", alpha = 0.8, position = position_stack(reverse = TRUE), fill = "#8C3F4D") +
  # facet_share(~gender, dir = "h",reverse_num = TRUE, scales = "free", switch = "both", strip.position = "top") + coord_flip()+
    labs(subtitle = "N = 553,262",
   caption = "Figure 1: Multimorbid patients per age group.", fill = "No of Conditions")+ 
  scale_color_manual(name = "", values = c(Male = "#3E606F", Female = "#8C3F4D"))+
  theme_classic() +
 theme(text = element_text(color = "#3A3F4A"),
    panel.grid.major.y = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    # axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.x = element_text(size = 12, color = "#5D646F", angle = 45,  hjust=1),
    # axis.text.y = element_blank(),
    # strip.text = element_text(color = "#5D646F", size = 18, face = "bold", hjust = 0.030),
    plot.background = element_rect(fill = "#EFF2F4"),
    # plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "bottom",
    legend.spacing = unit(0.1, "lines"),
    legend.text  = element_text(size = 8),
    legend.text.align = 0, 
    legend.direction = "horizontal"
    ) +
  guides(fill = guide_legend(nrow = 1))
```

```{r, include=FALSE}
mm_df5 <- inputTemp2 %>%
  group_by(rowId) %>%
  filter(n()>= 2) %>%
  group_by(condition) %>%
  summarise(N = n()) 

p3 <- mm_df5 %>%
  arrange(desc(N)) %>%
  slice_head(n =30) %>%
  mutate(condition = str_replace_all(condition, " / ", "/")) %>%
ggplot(aes(x = reorder(condition, -N), y = N)) +
  geom_bar(stat = "identity", width = 0.9, fill ="#56B4E9") +
  labs(x = "Condition",
   subtitle = "",
   caption = "Figure 3: Top 30 Multimorbid conditions.", 
   fill = "")+ 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
  # scale_x_discrete(labels = function(x) str_wrap(x, width = 40))+
  theme_classic() +
 theme(text = element_text(color = "#3A3F4A"),
    panel.grid.major.y = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    # axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.x = element_text(size = 10, color = "#5D646F", angle = 70,  hjust=1),
    # axis.text.y = element_blank(),
    # strip.text = element_text(color = "#5D646F", size = 18, face = "bold", hjust = 0.030),
    plot.background = element_rect(fill = "#EFF2F4"),
    # plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "bottom",
    legend.spacing = unit(0.1, "lines"),
    legend.text  = element_text(size = 8),
    legend.text.align = 0, 
    legend.direction = "horizontal"
    ) +
  guides(fill = guide_legend(nrow = 1)) 
```

Figure 1: Age at multimorbidity diagnosis.

```{r, echo=FALSE}
p1
```

Figure 2: Ag3 distribution according to the number of conditions diagnosed.

```{r, echo=FALSE}
p2
```

Figure 3: Top 30 Multimobid conditions

```{r, echo=FALSE}
p3
```

### Associative conditions

```{r, include=FALSE}
df_inputTemp <- as.data.frame(dplyr::select(inputTemp2, c(rowId, condition, timeId)))

mmDf <- df_inputTemp %>%
  group_by(rowId) %>%
  mutate(size = n()) %>%
  filter(size!= 1) %>%
  # arrange(desc(timeId)) %>%
  mutate(eventId = dense_rank(desc(timeId))) %>%
  arrange(rowId, eventId) %>%
  ungroup() 

mmDf <- as.data.frame(mmDf)
length(unique(mmDf$rowId))

# ARM
aprioriInput <- as(split(mmDf[,"condition"], mmDf[,"rowId"]), "transactions")

minSup = 10/populationWithMultimorbidity
res1 <- apriori(aprioriInput, parameter = list(support = minSup, confidence = 0, minlen= 2, maxlen = 5, maxtime = 0), control = list(verbose = TRUE))
res1
res2 <- res1[!is.redundant(res1)]
res2
```

```{r}
quality(res2) <- quality(res2) %>% 
  mutate(hyperConfidence = interestMeasure(res2, measure = "hyperConfidence", transactions = aprioriInput), 
         stdLift = interestMeasure(res2, "stdLift", transactions = aprioriInput), 
         chisq = interestMeasure(res2, "chiSquared", transactions = aprioriInput, significance = TRUE), 
         relativeRisk = interestMeasure(res2, "relativeRisk", transactions = aprioriInput), 
         phi = interestMeasure(res2, "phi", transactions = aprioriInput), 
         jaccard = interestMeasure(res2, "jaccard", transactions = aprioriInput), 
         adjusted_chisq = p.adjust(chisq, "fdr"),
         size = size(res2))



quality(res2) %>% 
  filter(lift > 1) %>%
  as_tibble()

quality(res2) %>%  
  filter(lift > 1) %>%
  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) 
quality(res2) %>%  
  filter(lift > 1) %>%
 summarise(across(where(is.numeric), ~ median(.x, na.rm = TRUE)))

quality(res2) %>%  
  filter(lift > 1 & relativeRisk >2) %>%
  as_tibble()

head(as(res2, "data.frame"))

res2df <- as(res2, "data.frame") %>%
  as_tibble() %>%
  filter(stdLift >= 0.2)

resTest <- as(subset(res2, subset = lift > 1 & relativeRisk>1 & count > 1000 & stdLift > 0.2 & confidence > 0.4), "data.frame") %>% as_tibble()

res2 <- subset(res2, subset = stdLift >0.2)
res2

res3 <- res2[is.maximal(res2)]
# is.superset(res3, res3)
# res3[is.superset(res3)]
res3
# resClosed <- res2[is.closed(res2)]
quality(res3) <- quality(res3) %>%
  mutate(hyperConfidence = interestMeasure(res3, measure = "hyperConfidence", transactions = aprioriInput), 
         stdLift = interestMeasure(res3, "stdLift", transactions = aprioriInput), 
         chisq = interestMeasure(res3, "chiSquared", transactions = aprioriInput, significance = TRUE), 
         adjusted_chisq = p.adjust(chisq, "fdr"),
         size = size(res3))

res3df <- as(res3, "data.frame")

```

```{r, include=FALSE}
quality(res2) <- quality(res2) %>%
  mutate(hyperConfidence = interestMeasure(res2, measure = "hyperConfidence", transactions = aprioriInput), 
         stdLift = interestMeasure(res2, "stdLift", transactions = aprioriInput), 
         chisq = interestMeasure(res2, "chiSquared", transactions = aprioriInput, significance = TRUE), 
         adjusted_chisq = p.adjust(chisq, "fdr"),
         size = size(res2))

res2df <- as(res2, "data.frame") %>%
  as_tibble() %>%
  filter(stdLift >= 0.2)

res2 <- subset(res2, subset = stdLift >0.2)
res2

res3 <- res2[is.maximal(res2)]
# is.superset(res3, res3)
# res3[is.superset(res3)]
res3
# resClosed <- res2[is.closed(res2)]
quality(res3) <- quality(res3) %>%
  mutate(hyperConfidence = interestMeasure(res3, measure = "hyperConfidence", transactions = aprioriInput), 
         stdLift = interestMeasure(res3, "stdLift", transactions = aprioriInput), 
         chisq = interestMeasure(res3, "chiSquared", transactions = aprioriInput, significance = TRUE), 
         adjusted_chisq = p.adjust(chisq, "fdr"),
         size = size(res3))

res3df <- as(res3, "data.frame")

res3
res4 <- subset(res3, subset = stdLift >= 0.2)
res4

res5 <- generatingItemsets(res4)
res5
res6 <- unique(res5)
res6
# res4df <- as(res4, "data.frame") %>%
#   as_tibble()
# 
# res6df <- as(res6, "data.frame") %>%
#   as_tibble() %>%
#   mutate(size = size(res6))
# # size(res6)


res7 <- subset(res4, subset = confidence >=0.8)

# res7df <- as(res7, "data.frame") %>%
#   as_tibble()
```

```{r, include=FALSE}
# res4df %>%
#   group_by(size) %>%
#   count()
# res6df %>%
#   group_by(size) %>%
#   count()
# res6df %>%
#   arrange(size) %>%
#   slice_head(n=9L)
# res7df%>%
#   group_by(size) %>%
#   count()
# res7df %>%
#   arrange(size) %>%
#   slice_head(n=9L)
# subset(res4, subset = x %in% res6) 
# plot(res7)

```

```{r, include=FALSE}
# res6df %>% 
#   ggplot(aes(x = count, y = confidence, fill = stdLift)) +
#   geom_point()
# 
# res4df %>% 
#   ggplot(aes(x = count, y = confidence, color = stdLift)) +
#   geom_point()
# 
# res2df %>% 
#   ggplot(aes(x = count, y = confidence, fill = stdLift)) +
#   geom_point()
# 
# p <- res3df %>% 
#   ggplot(aes(x = support, y = confidence, text = rules, size = size)) +
#   geom_point(alpha = 0.5, aes(color = stdLift))
# p
# # ggplotly(p, tooltip = "text")
# 
# res3df %>% arrange(desc(stdLift)) %>% head()
# res3df %>% arrange(desc(support)) %>% head()
# 
# res3df %>%
#   ggplot(aes(x = support, y = stdLift, color = chisq, size = confidence)) +
#   geom_point()
# 
# res3df %>% filter(adjusted_chisq < 0.00000005) %>% count()
# unique(generatingItemsets(res2))
# res6.5 <- res6
# quality(res6.5) <- quality(res6.5) %>%
#   mutate(allConfidence = interestMeasure(res6.5, measure = "allConfidence", transactions = aprioriInput), 
#          stdLift = interestMeasure(res6.5, "stdLift", transactions = aprioriInput))


# class(res2)
# addAdditionalMeasures(res6.5, aprioriInput)
res3g <- associations2igraph(res6, associationsAsNodes = F)
res3g_undi <- as.undirected(res3g,  mode = "collapse", edge.attr.comb = "sum")
res3g_undi
# crc1 <- layout_in_circle(res3g_undi)
# plot(res3g_undi, layout = crc1)
# crc <- layout_as_star(res3g_undi, center = V(res3g_undi)[3])
# plot(res3g_undi, layout = crc)
# plot(res3g)

set.seed(1)
# ggraph(res3g, 'circlepack', weight = size) + 
#   geom_edge_link() + 
#   geom_node_point(aes(colour = support)) +
#   coord_fixed()

vertexWeight <- tibble(freq = itemFrequency(aprioriInput), nam = names(itemFrequency(aprioriInput))) %>%
  mutate(count = freq*populationWithMultimorbidity)

# vertexWeight <- vertexWeight %>%
#   inner_join(concepts %>% 
#                select(condition, Category) %>%
#                distinct(), by = join_by("nam" == "condition")
#              )

# vertexWeight <- vertexWeight %>%
#   inner_join(concepts %>% 
#                select(conceptName, Category) %>%
#                distinct(), by = join_by("nam" == "conceptName")
#              )

vertexWeight <- vertexWeight %>%
  inner_join(concepts %>% 
               select(condition, Category) %>%
               distinct(condition), by = join_by("nam" == "condition")
             )

V(res3g_undi)$weightFreq = vertexWeight %>% filter(nam %in% V(res3g_undi)$label) %>% pull(freq)
V(res3g_undi)$weightCount = vertexWeight %>% filter(nam %in% V(res3g_undi)$label) %>% pull(count)
# V(res3g_undi)$groupCategory = vertexWeight %>% filter(nam %in% V(res3g_undi)$label) %>% pull(Category)
V(res3g_undi)$groupCategory = "oneCat"
edgeAttrs <- quality(res6)
E(res3g_undi)$
res3g_undi
pg <- ggraph(res3g_undi, "circle") +
  # geom_edge_arc0(position = "jitter", aes(edge_width = support), color = "grey", alpha = 0.5,
  #                show.legend = FALSE) +
  geom_edge_fan0(position = "jitter", aes(edge_width = support), color = "grey", alpha = 0.5,
                 show.legend = FALSE) +
  geom_node_point(position = "jitter", aes(size = weightCount, color = groupCategory), show.legend = FALSE)+
  # geom_node_text(aes(label = label, angle = node_angle(x, y)), hjust = -0.1)+
  geom_node_text(aes(label = str_wrap(stringr::str_remove_all(string = label, pattern = " / "), width = 15)), check_overlap = TRUE,  size = 3, repel = TRUE, seed = 1) +
  # geom_text(label.size = NA)+
  # geom_node_label(aes(label = label), color = "blue") +
  scale_x_continuous(expand = expansion(mult = 0.1)) +
   # scale_x_continuous(labels = function(x) str_wrap(x, whitespace_only = TRUE), expand = expansion(mult = 0.1)) + 
   # scale_label_size_continuous(guide = guide_axis(check.overlap = TRUE), labels = function(x) str_wrap(x, width = 20), size) +
  ggraph::theme_graph(strip_text_size = 2)

# ggsave(plot = p, filename = "results/OHDSI_Europe/graph.jpeg")
```

Figure 4: Graph indicating 2 way association between conditions grouped by category.

```{r, echo=FALSE}
pg
```

```{r, include=FALSE}
minSup = 10/populationWithMultimorbidity
res1 <- apriori(aprioriInput, parameter = list(support = minSup, confidence = 0, minlen= 2, maxlen = 25, maxtime = 0), control = list(verbose = TRUE))
res1
res2 <- res1[!is.redundant(res1)]

quality(res2) <- quality(res2) %>%
  mutate(hyperConfidence = interestMeasure(res2, measure = "hyperConfidence", transactions = aprioriInput), 
         stdLift = interestMeasure(res2, "stdLift", transactions = aprioriInput), 
         chisq = interestMeasure(res2, "chiSquared", transactions = aprioriInput, significance = TRUE), 
         adjusted_chisq = p.adjust(chisq, "fdr"),
         size = size(res2))
res2
res2 <- subset(res2, subset = stdLift >0.2)
res2
res2 <- subset(res2, subset = confidence >= 0.5)
res2
res2[!is.redundant(res2)]

as(res2, "data.frame")%>% 
  ggplot(aes(x = support, y = confidence, text = rules, size = size)) +
  geom_point(alpha = 0.5, aes(color = stdLift))
# p


res2Df <- as(unique(generatingItemsets(res2)), "data.frame")
res1Df <- as(res2, "data.frame")

test <- res1Df %>%
  mutate(newRules = stringr::str_replace_all(rules, pattern = "\\} => \\{", ",")) %>%
  filter(newRules %in% res2Df$items)

# # find all subsets in our rule set using is.subset. Note that we have already sorted our rules by descending lift.
# rules.subset<-is.subset(res2)
# # throw away the bottom half, including the diagonal, of our identity matrix
# rules.subset[lower.tri(rules.subset,diag=T)]<-NA
# 
# # find all matching subset rules
# redundant<-colSums(rules.subset,na.rm = T)>=1
# # prune matching subset rules
# rules.pruned<-rules[!redundant]

tab1 <-test %>% 
  group_by(size) %>%
  count() %>%
  ungroup() %>%
  mutate(size = as.character(size)) %>%
  do(bind_rows(., tibble(size = "Total", n = nrow(test)))) %>%
  gt()

tab1

itemFrequencyPlot(aprioriInput)
tibble(name = names(itemFrequency(aprioriInput)), 
       freq = itemFrequency(aprioriInput), 
       pop = itemFrequency(aprioriInput)*populationWithMultimorbidity) %>%
  slice_max(order_by = freq, n = 20) %>%
  ggplot(aes(x = name, y = freq))+
  geom_bar(stat = "identity", alpha = 0.2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



freqs <- tibble(name = names(itemFrequency(aprioriInput)), 
       freq = itemFrequency(aprioriInput), 
       pop = itemFrequency(aprioriInput)*populationWithMultimorbidity) %>%
  arrange(desc(pop))
freqs[1:12,] %>% pull(name)
label_data <- freqs %>%
  mutate(newLabs = abbreviate(name))

number_of_bar <-12
label_data <- label_data[1:12,] %>% 
  mutate(rownum = row_number())
angle <-  90 - 360 * (label_data$rownum-0.5) /12 

label_data$hjust<-ifelse( angle < -90, 1, 0)
freqs %>%
  slice_head(n = 12) %>%
  ggplot(aes(x = reorder(name, -pop, sum), y = pop)) +
  geom_bar(stat = "identity", fill = "lightblue4") +
  # coord_polar(start = 0) + 
   geom_text(data=label_data, aes(x=name, y=pop+10, label=str_wrap(name, 10), hjust=0.5, vjust =-0.1),  color="black", fontface="bold",alpha=0.6, size=3#label_data$hjust,inherit.aes = FALSE 
                )  +
  theme_classic() +
  scale_y_continuous(limits = c(0, 230000), labels = function(x) if_else(x == 0, "0", paste0(as.character(x/1000), "K"))) +
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
        # axis.text = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(face = "bold"),
    # axis.ticks.y = element_blank(),
    axis.title = element_blank(), 
    # panel.grid = element_blank()
    # plot.margin = unit(rep(1,4), "cm")
    panel.grid.major.y = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A")
    ) 

hjustVec <- c(1, rep(-0.1, 11))
vjustVec <- c(1.5, rep(0.5, 11))
freqs %>%
  slice_head(n = 12) %>%
  ggplot(aes(x = reorder(name, pop, sum), y = pop)) +
  geom_col( fill = "darkblue", alpha = 0.7) +
  # geom_bar(data = freqs[13:24,], aes(x = reorder(name, pop, sum), y = pop),stat = "identity", position = "dodge", fill = "darkblue", alpha = 0.3) +
  # coord_polar(start = 0) + 
   geom_text(data=label_data, aes(x=name, y=pop, label=name),  color="black", fontface="bold", size=5, inherit.aes = TRUE, hjust = 0.4)  +
  coord_flip() +
  theme_classic() + 
  scale_y_continuous(limits = c(0, 230000), labels = function(x) if_else(x == 0, "0", paste0(as.character(x/1000), "K"))) +
  labs(title = "Figure 1: Top 10 multimorbid conditions.")+
  theme(#axis.text.x = element_text(angle = 45, hjust = 1),
        # axis.text = element_blank(),
     plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(face = "bold"),
    # axis.ticks.y = element_blank(),
    axis.title = element_blank(), 
    panel.grid.major.y = element_line(linetype = "dotted", size = 0.01, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"), 
    plot.background = element_rect(fill = "#EFF2F4")
    # panel.grid = element_blank()
    # plot.margin = unit(rep(1,4), "cm")
    ) 

# ggsave(filename = "results/OHDSI_Europe/figure4_top10MultimorbidConditions.jpeg", width = 10, height =10)


 theme(text = element_text(color = "#3A3F4A"),
    panel.grid.major.y = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    # axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.x = element_text(size = 10, color = "#5D646F", angle = 70,  hjust=1),
    # axis.text.y = element_blank(),
    # strip.text = element_text(color = "#5D646F", size = 18, face = "bold", hjust = 0.030),
    plot.background = element_rect(fill = "#EFF2F4"),
    # plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "bottom",
    legend.spacing = unit(0.1, "lines"),
    legend.text  = element_text(size = 8),
    legend.text.align = 0, 
    legend.direction = "horizontal"
    ) +
  guides(fill = guide_legend(nrow = 1)) 
```

```{r, include=FALSE}
res2g <- associations2igraph(res2, associationsAsNodes = F)
res2g_undi <- as.undirected(res2g,  mode = "collapse", edge.attr.comb = "sum")
# res2g_nd3 <- networkD3::igraph_to_networkD3(res2g_undi)
# str(res2g_nd3$links)
# adjacencyMatrix <- with(res2g_nd3$links, table(source, target))

V(res2g_undi)$name <- abbreviate(V(res2g_undi)$label)
adjacencyMatrix2 <- igraph::as_adjacency_matrix(res2g_undi, attr = "count", sparse = FALSE, names = TRUE)
rownames(adjacencyMatrix2)
colnames(adjacencyMatrix2)
plotDf <- reshape2::melt(adjacencyMatrix2)
names(plotDf)
# abbreviate(V(res2g_undi)$label)
hm <- ggplot(plotDf, aes(Var1, Var2, fill=value )) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE)+ 
  theme_bw() +
    theme(text = element_text(color = "#3A3F4A"),
    # panel.grid.major.y = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    panel.grid.minor = element_blank(),
    # panel.grid.major.x = element_line(linetype = "dotted", size = 0.1, color = "#3A3F4A"),
    # axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 16, margin = margin(b = 10), hjust = 0.030),
    plot.subtitle = element_text(size = 12, hjust = 0.030),
    plot.caption = element_text(size = 8, color = "#5D646F", hjust = 0.0),
    axis.text.x = element_text(size = 9, color = "black", angle = 45,  hjust=1),
    axis.text.y = element_text(size = 9, color = "black"),
    # axis.text.y = element_blank(),
    # strip.text = element_text(color = "#5D646F", size = 18, face = "bold", hjust = 0.030),
    plot.background = element_rect(fill = "#EFF2F4"),
    # plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.position = "right",
    # legend.spacing = unit(0.1, "lines"),
    # legend.text  = element_text(size = 8),
    legend.text.align = 0, 
    # legend.direction = "horizontal"
    ) 

```

Figure 5: Heatmap of non redundant rules, with minsup \> 0.000018, confidence \> 0.5 and stdLift \> 0.2.

```{r, echo=FALSE,  fig.width=11, fig.height=11}
# library(grid)
# grid.newpage()
hm
```

(This plot will be improved by adding a legend with the abbreviated names).

```{r, echo=FALSE, warning=FALSE}
# L <- tibble(name =  V(res2g_undi)$label, abr = abbreviate(V(res2g_undi)$label))
#  grid.newpage()
# grid.arrange(
#   hm,
#   tableGrob(L),
#     ncol = 2,
#   widths = c(1.5, 1)
# )
```

Table 5: Set of association rules with minsup 0.000018, minconf 0.5 and stdLift \> 0.2 `r tab1`

```{r, include=FALSE}
top2 <- test %>%
  filter(size ==2) %>%
  arrange(desc(stdLift)) %>%
  slice_head(n = 10) %>%
  select(rules, support, count, confidence, stdLift)

top3 <- test %>%
  filter(size ==3 ) %>%
  arrange(desc(stdLift)) %>%
  slice_head(n = 10) %>%
  select(rules, support, count, confidence, stdLift)

top4 <- test %>%
  filter(size ==4) %>%
  arrange(desc(stdLift)) %>%
  slice_head(n = 10)  %>%
  select(rules, support, count, confidence, stdLift)

top5 <- test %>%
  filter(size ==5) %>%
  arrange(desc(stdLift)) %>%
  slice_head(n = 10)%>%
  select(rules, support, count, confidence, stdLift)

top6 <- test %>%
  filter(size ==6) %>%
  arrange(desc(stdLift)) %>%
  slice_head(n = 10)%>%
  select(rules, support, count, confidence, stdLift)
```

Table 6: Top rules based on stdLift stratified by size **Size 2**. `r top2 %>% gt()` **Size 3**.\
`r top3 %>% gt()` **Size 4**. `r top4 %>% gt()` **Size 5**. `r top5 %>% gt()` **Size 6** `r top6 %>% gt()`

# References
